Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE1)

### 2014-04-10 WSSSPE1_report/WSSSPE1_summary_report.tex
add my affiliation
starting to work on cross-cutting Career Track section - in progress

\thanks{{}$^{(1)}$ National Science Foundation, Arlington, VA, USA; Computation Institute, University of Chicago \& Argonne National Laboratory, Chicago, IL, USA; \url{d.katz@ieee.org}}
%
\thanks{{}$^{(2)}$ National Evolutionary Synthesis Center (NESCent),
  Durham, NC, USA; \url{hlapp@nescent.org}}
%
\thanks{{}$^{(3)}$ Argonne National Laboratory, Chicago, IL, USA; \url{ketan@anl.gov}}
%
\thanks{{}$^{(4)}$ xxx}
%
\thanks{{}$^{(5)}$ Columbia Astrophysics Laboratory, Columbia University, New
York, NY, USA; \url{matthewturk@gmail.com}}
%
\thanks{{}$^{(7)}$ xxx}
%
\thanks{{}$^{(8)}$ Leader, Research Software Development Team, University College London; \url {j.hetherington@ucl.ac.uk}}
%
\thanks{{}$^{(9)}$ xxx}
%
\thanks{{}$^{(10)}$ NORC at the University of Chicago and the Illinois Institute of Technology, Chicago, IL, USA; (\url{sctchoi@uchicago.edu}}
% 
\thanks{{}$^{(11)}$ xxx}
%
\thanks{{}$^{(12)}$ xxx}
%
\thanks{{}$^{(13)}$ xxx}
%
\thanks{{}$^{(14)}$ xxx}

@@

He then presented arguments on directly valuing software as
a research output alongside papers, a common discussion within this field.
An interesting reference provided by him
is~\cite{peer-review-code}, which explores involving software
engineers in the review process of scientific code.

@@

Recruiting and/or retaining personnel in this area is hard,
with one of the major reasons being no long-term career paths (especially when
compared with industry). How should software development be balanced with
research? It is apparent that things are beginning to improve, such as
incentives for software development in the form of altmetrics, tenure committee
consideration, and NSF ``products'' vs ``publications.'' It was noted it can be very
difficult to measure where people are using small bits of your code.

@@

How to ensure software is sustainable by ensuring there are career
paths for developers.

>>>>

\thanks{{}$^{(1)}$ National Science Foundation, Arlington, VA, USA; Computation Institute, University of Chicago \& Argonne National Laboratory, Chicago, IL, USA; \url{d.katz@ieee.org}}
%
\thanks{{}$^{(2)}$ National Evolutionary Synthesis Center (NESCent),
  Durham, NC, USA; \url{hlapp@nescent.org}}
%
\thanks{{}$^{(3)}$ Argonne National Laboratory, Chicago, IL, USA; \url{ketan@anl.gov}}
%
\thanks{{}$^{(4)}$ Center for Computation and Technology, Louisiana State University, Baton Rouge, LA, USA; \url{knarf@cct.lsu.edu}}
%
\thanks{{}$^{(5)}$ Columbia Astrophysics Laboratory, Columbia University, New
York, NY, USA; \url{matthewturk@gmail.com}}
%
\thanks{{}$^{(7)}$ xxx}
%
\thanks{{}$^{(8)}$ Leader, Research Software Development Team, University College London; \url {j.hetherington@ucl.ac.uk}}
%
\thanks{{}$^{(9)}$ xxx}
%
\thanks{{}$^{(10)}$ NORC at the University of Chicago and the Illinois Institute of Technology, Chicago, IL, USA; (\url{sctchoi@uchicago.edu}}
% 
\thanks{{}$^{(11)}$ xxx}
%
\thanks{{}$^{(12)}$ xxx}
%
\thanks{{}$^{(13)}$ xxx}
%
\thanks{{}$^{(14)}$ xxx}

@@

He then presented arguments on directly valuing software as
a research output alongside papers, a common discussion within this field.
He mentioned an interesting exploration of involving software
engineers in the review process of scientific code~\cite{peer-review-code}.

@@

Recruiting and/or retaining personnel in this area is hard,
with one of the major reasons being no long-term career paths (especially when
compared with industry). This is further discussed in \S\ref{sec:career-tracks}.

%moved to cross-cutting issues
%
%How should software development be balanced with
%research? It is apparent that things are beginning to improve, such as
%incentives for software development in the form of altmetrics, tenure committee
%consideration, and NSF ``products'' vs ``publications.'' It was noted it can be very
%difficult to measure where people are using small bits of your code.

@@

Career track issues also came up repeatedly, starting in the the first keynote (\S\ref{sec:keynote1}), where Phil Bourne used the term ``the Google Bus'' to describe the issue of staff in his lab in San Diego who had chosen to leave science and work for Google (but since they didn't want to physically move, they take the Google bus to the Google office each day.)

The lack of career paths also came up in the Developing and Supporting Software discussion (\S\ref{sec:devel}), as an issue that led to difficulties in established projects retaining as well as recruiting staff. The issue there was that for software to be sustainable, there need to be career paths for software developers.  There was also some discussion about how should software development be balanced with research, with the idea that faculty are hired based on their research qualifications, not on their software development qualification, and as such, specializing in software development was likely to not lead to a career in science at a university, though clearly, national laboratories are a different case.

In terms of possible solutions, the Developing and Supporting Software discussion (\S\ref{sec:devel}) said that things are beginning to improve, for example, with incentives for software development in the form of altmetrics, tenure committee consideration, and NSF's change in its biosketches of asking for proposer's ``products'' rather than their ``publications,'' as had been done previously.
However, it was also noted it can be very difficult to measure where people are using small bits of your code.

Open questions:
